Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-09-22 Paper/method.tex

In this section, we present a generalized framework for computing the
active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
problem whereby the 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Specifically, we employed two strategies --
a perturbation-based strategy that involves computation of model
gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a regression-based approach 
that involves a linear approximation of
the model output to approximate the gradient. 
The two approaches are based on Algorithms~1.1 and 1.2 in~\cite{Constantine:2015} and
essentially differ in the way the gradient is estimated. They were implemented in an
iterative manner to ensure computational efficiency by avoiding 
unnecessary model evaluations once a converged active subspace
was achieved.

@@

As discussed earlier, gradient estimation using finite difference in the 
perturbation-based strategy requires additional model evaluations at the
neighboring points in the input domain. 
Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
model evaluations are needed. On the other hand, the regression-based
strategy uses the available set of $N$ model evaluations to approximate
the gradient. The computational effort is therefore 
reduced by a factor $(d+1)$ in this case. The specific sequence of steps
for computing the active subspace is discussed as follows.

>>>>

In this section, we present an efficient framework for computing the
active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
problem whereby the 
pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
Table~\ref{tab:kinetics} is considered to be uniformly
distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ is the nominal
estimate provided in~\cite{Yetter:1991}.
Specifically, we employed two strategies: 
a perturbation-based strategy that involves computation of model
gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
in~\eqref{eq:chat}, and a regression-based approach 
that involves a local linear approximation of
the model output to approximate the gradient. 
The two approaches are based on Algorithms~1.1 and 1.2 in~\cite{Constantine:2015} and
essentially differ in the way the gradient is estimated. They were implemented in an
iterative manner to ensure computational efficiency by avoiding 
unnecessary model evaluations once a converged active subspace
was achieved.

@@

As discussed earlier, gradient estimation using finite differences in the 
perturbation-based strategy requires additional model evaluations at the
neighboring points in the input domain. 
Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
model evaluations are needed. On the other hand, the regression-based
strategy uses the available set of $N$ model evaluations to approximate
the gradient. The computational effort is therefore 
reduced by a factor $(d+1)$ in this case. The specific sequence of steps
for computing the active subspace is discussed as follows.
